#Reproducible Research: Peer Assessment 1
Author: Sainath Kovelamudi


## Loading and preprocessing the data

* Loading the data
```{r echo = TRUE}
activity <- read.csv("activity.csv")
summary(activity)
```

* Processing the data
```{r echo = TRUE}
activity$date <- as.Date(activity$date,"%Y-%m-%d")
```


## What is mean total number of steps taken per day?
* calculating the total number of steps taken per day
```{r echo =TRUE}
totalSteps <- aggregate(steps~date,data=activity,sum,na.rm=TRUE)
head(totalSteps)
```

* histogram for total number of steps taken per day
```{r echo = TRUE}
hist(totalSteps$steps,
     breaks=nrow(totalSteps),col="blue",
     xlab="Total number of steps per day",
     ylab="Frequency",
     main="Histogram")
```

*The MEAN of the total number of steps taken per day : `r mean(totalSteps$steps)`.
*The MEDIAN of the total number of steps taken per day : `r median(totalSteps$steps)`.

## What is the average daily activity pattern?
*Average number of steps taken in a time interval
```{r echo=TRUE}
avgSteps <- aggregate(steps~interval,data=activity,mean,na.rm=TRUE)
plot(steps~interval,data=avgSteps,type="l",
      xlab="Time intervals",
      ylab="Average number of steps",
      col="blue")
```
*The maximum number of steps are taken in the interval: `r avgSteps[which.max(avgSteps$steps),]$interval`.

## Imputing missing values
* To see total number of missing values in the dataset
```{r echo=TRUE}
summary(is.na(activity))
```
*imputing the missing steps value with the mean of the number of steps value and creating a new dataset with the missing values filled

```{r echo=TRUE}
meanSteps<-mean(activity[!is.na(activity$steps),]$steps)
activity1<-activity
activity1[is.na(activity1$steps),]$steps <- meanSteps

``` 
```{r echo =TRUE}
totalSteps <- aggregate(steps~date,data=activity1,sum)
```
* Histogram of total number of steps taken each day
```{r echo = TRUE}
hist(totalSteps$steps,breaks=nrow(totalSteps),col="blue",
     xlab="Total number of steps per day",
     ylab="Frequency",
     main="Histogram") 
```
*The MEAN of the total number of steps taken per day : `r mean(totalSteps$steps)`.
*The MEDIAN of the total number of steps taken per day : `r median(totalSteps$steps)`.

## Are there differences in activity patterns between weekdays and weekends?
* creating a new factor variable with two levels 'weekday' and 'weekend'
```{r echo = TRUE}
day <- weekdays(activity1$date)
activity1$day <- ifelse(day == "Saturday"| day=="Sunday","weekend","weekday")
head(activity1)
```

* Making a panel plot of average number of steps taken in a interval, averaged across all weekdays or weekdays.

```{r echo = TRUE}
avgSteps1 <- aggregate(activity1$steps,
                      by=list(activity1$interval,activity1$day),mean)
names(avgSteps1) <- c("interval","day","steps")
library(lattice)
xyplot(steps~interval | day, data=avgSteps1, type="l",layout=c(1,2),
       xlab="Interval",ylab="Number of steps",
       main="Activity Patterns on weekends and weekdays",col="blue")
```
* The activity pattern for weekday has a big spike (peak value = 200) at around the interval value 800 which is not the same in the weekend pattern. The weekend pattern has more number of steps for most other intervals.